---
title: 'Reproducible Research: Peer Assignment 2'
author: "Yongyang Huo"
date: "24 May 2015"
output: html_document
---

## Weather Events Impact on Public Health and Economy Analysis


### Synonpsis  
In this report, based on the storm data collected by U.S. National Oceanic and Atmospheric Administration's (NOAA), we analysis the data and find out the relationship between weather event and population health, economic respectively. the following questions is being marked with answers:
        1.	Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?
	Excessive heat and tornado are most harmful 
	2. 	Across the United States, which types of events have the greatest economic consequences?
	Flood, drought, and hurricane/typhoon have the greatest economic 				consequences

###  Config
```{r}
echo = TRUE  
options(scipen = 1)  
library(R.utils)
library(ggplot2)
library(plyr)
require(gridExtra)
```

## Loading and preprocessing the data
```{r}
filedir <- "./data/repdata-data-StormData.csv"
#getwd()
stormData <- read.csv(filedir, header = TRUE)
#Generate Year for further filtering
stormData$year <- as.numeric(format(as.Date(stormData$BGN_DATE, format = "%m/%d/%Y %H:%M:%S"), "%Y"))
hist(stormData$year, breaks = 30)
```


```{r}
storm <- stormData[stormData$year >= 1995, ]
```

#### Weather Event impact on Public Health
```{r}
sortHelper <- function(fieldName, top = 15, dataset = stormData) {
    index <- which(colnames(dataset) == fieldName)
    field <- aggregate(dataset[, index], by = list(dataset$EVTYPE), FUN = "sum")
    names(field) <- c("EVTYPE", fieldName)
    field <- arrange(field, field[, 2], decreasing = T)
    field <- head(field, n = top)
    field <- within(field, EVTYPE <- factor(x = EVTYPE, levels = field$EVTYPE))
    return(field)
}

fatalities <- sortHelper("FATALITIES", dataset = storm)
injuries <- sortHelper("INJURIES", dataset = storm)
```

#### Weather Event impact on Economy

```{r}
convertHelper <- function(dataset = storm, fieldName, newFieldName) {
    totalLen <- dim(dataset)[2]
    index <- which(colnames(dataset) == fieldName)
    dataset[, index] <- as.character(dataset[, index])
    logic <- !is.na(toupper(dataset[, index]))
    dataset[logic & toupper(dataset[, index]) == "B", index] <- "9"
    dataset[logic & toupper(dataset[, index]) == "M", index] <- "6"
    dataset[logic & toupper(dataset[, index]) == "K", index] <- "3"
    dataset[logic & toupper(dataset[, index]) == "H", index] <- "2"
    dataset[logic & toupper(dataset[, index]) == "", index] <- "0"
    dataset[, index] <- as.numeric(dataset[, index])
    dataset[is.na(dataset[, index]), index] <- 0
    dataset <- cbind(dataset, dataset[, index - 1] * 10^dataset[, index])
    names(dataset)[totalLen + 1] <- newFieldName
    return(dataset)
}

storm <- convertHelper(storm, "PROPDMGEXP", "propertyDamage")
storm <- convertHelper(storm, "CROPDMGEXP", "cropDamage")
names(storm)
options(scipen=999)
property <- sortHelper("propertyDamage", dataset = storm)
crop <- sortHelper("cropDamage", dataset = storm)
```


### Results
From these data, we found that excessive heat and tornado are most harmful with respect to population health